---
title: "A3: Incarceration"
author: "Chufeng Wang"
date: "08/04/2023"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(dplyr)
library(purrr)
library(maps)
```

### Introduction

The United States has the highest incarceration rate in the world. Disparities across different racial and ethnic groups in the US justice system have long been a subject of study and concern. Understanding these disparities is crucial to addressing the root causes of inequality in the system. The reason for selecting these variables is that they give us the data needed to analyze racial and ethnic disparities in incarceration rates in Washington. They allow us to examine the issue over time, between different groups, and across different regions in the state.

```{r summary, echo=FALSE, message=FALSE}
source("summary.R")
```

### Summary Information

The resultant variables contain key statistics for different racial and ethnic groups in Washington's jails for the year 2018. The average jail population rate of native, latinx, and appi are `r avg_jail_2018_native` , `r avg_jail_2018_latinx` and `r avg_jail_2018_aapi` in 2018. Moreover, the maximum jail population rate of native, latix, aND appi are `r max_jail_rate_native` , `r max_jail_rate_latinx` and `r max_jail_rate_aapi`.



### The Dataset

Who collected the data?\
The Vera Institute of Justice

How was the data collected or generated?\
The Vera organization gather the data from numerous data sources including: The Bureau of Justice Statistics (BJS) data collections, The Census of Jails (COJ), The Annual Survey of Jails (ASJ), The Centers for Disease Control and Prevention (CDC), The U.S. Department of Agriculture (USDA).

Why was the data collected?\
The data on racial disparities in incarceration was collected to identify and understand systemic racial biases within the U.S. criminal legal system. Analyzing this data helps in pinpointing trends that indicate discrimination, particularly against communities of color. By highlighting areas of inequity, it enables law enforcement agencies to adjust practices and fosters public awareness of systemic racism. The data serves as a crucial tool for promoting fairness, accountability, and human rights within the legal system.

How many observations (rows) are in your data?\
There are 1131 rows.

How many features (columns) are in the data?\
There are 23 columns.

What, if any, ethical questions or questions of power do you need to consider when working with this data?\
Does the data include personal or sensitive information that could be traced back to individuals? . Who collected the data, who has access to it, and who "owns" it? These questions of power can influence how the data is used, by whom, and for what purposes. How might the data be misinterpreted or misused, especially in a way that further marginalizes or stereotypes certain racial or social groups?

What are possible limitations or problems with this data? (at least 200 words)

The temporal constraints of the data might not capture recent changes in policies and social dynamics, limiting its current relevance. For example, the data only goes up to a certain date in 2018, it may not reflect recent changes in policies, social attitudes, or other factors affecting incarceration rates. Geographical inconsistencies and variations in racial demographics across counties can lead to potentially skew findings. A predominantly White county might show low Black incarceration rates, not because of fair practices but due to low Black population numbers.The lack of distinction between types of crimes and absence of contextual socioeconomic information may hinder a clear understanding of incarceration. Additionally, accessibility and interpretation challenges might limit the dataset's usability for various stakeholders.

### Trends Over Time Chart

The chart shows the number of different population groups been incarcerated changes since 1990 to 2018 in Washington and answer the question "Do native people different with other people in incarceration?"

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, code = readLines("chart1_example.R")}
```

### Variable Comparison Chart

The chart visualize and analyze the relationship between Native prison and jail population rates. It gives a better understanding of the complexities of the criminal justice system as it pertains to the Native community in Washington.

```{r, echo = FALSE, code = readLines("chart2_example.R")}
```

### Map

The resulting map visualizes how the Native jail rate is distributed across Washington's counties. This can provide insights into regional differences or trends in the Native jail rate, enabling policymakers to better understand and respond to the underlying factors driving these differences.

```{r, echo = FALSE, code = readLines("map_example.R")}
```
